\section{\huge Methods}
\subsection{Data Collection}
Complete genomes from NCBI RefSeq are downloaded and the \ac{crsp}db (along with a python script) is used to annotate genera as being mixed (containing strains with and without \ac{crspc} systems) or Non-\ac{crsp} (containing no strains with a \ac{crspc} system)\citep{crispdb}.
\ac{crsp} annotations of \ac{cas},Cfp proteins from NCBI and the \ac{crsp}one tool from Zhang and Ye will also be used to assess the presence of \ac{crsp} systems\citep{ineqcas}.
\subsection{Gene Presence/Absence Matrix}
In order to use the program markophylo to estimate insertion and deletion rates, a \ac{pa} matrix and a phylogenetic species tree are required.
First any genes classified as \ac{mge}s (from NCBI annotations) are removed.
Next genes are grouped into families by reciprocal BLAST hits and single link clustering.
Genes not placed in any family this way are BLASTed against the NCBI NR database to check if they are valid genes, if they are they are consider their own family.
The \ac{pa} matrix is constructed as follows, for each \ac{otu} a binary vector is created, where each entry represent a gene family and a 1 indicate that that \ac{otu} contains 1 gene in that family.
This is repeated for all \ac{otu}, creating a $G \times O$ binary matrix, where $G$ is the total number of gene families and $O$ is the number \ac{otu}s.\par
There are many ways to construct a species tree, but for this project the tree will be constructed using genes from gene families present in all \ac{otu}s being considered, using Bayesian methods, as implemented in the program MrBayes.
\subsection{Makophylo Rate Estimations}
Given a species tree and a gene family \ac{pa} matrix for the \ac{otu}s of the species tree the R package \textit{markophylo} can provide gene insertion and gene deletion rate estimates\citep{marko}.
The presence or absence of gene families are considered 2 discrete states, for which a $(2\times 2)$ transition rate matrix (of a \ac{mc} model) can be estimated using maximum likelihood techniques.
This values in this estimated transition matrix are the insertion rate (transition probability of gene absence $\to$ presence)and deletion rate (transition probability of gene presence $\to$ absence)\citep{marko}.
\subsection{Network Construction}
Quartet decomposition is method by which \ac{hgt} events can be identified using a set of gene trees and a species tree.
Given a tree $T$ a quartet is a subtree contain 4 of the leaf nodes in $T$, meaning that for a tree with $N$ leaf nodes (or \ac{otu}s) there are $\binom{N}{4}$ unique quartets in that tree.
A quartet $Q$ is considered consistent with a tree if $Q = T|Le(Q)$ where $T|Le(Q)$ is the tree obtained by suppressing all degree-two nodes in $T[X]$ and $T[X]$ is the minimal subtree of T with all nodes in $X$, which is a leaf set of $T$\citep{hide}.
To calculate the weight of an edge for the network, given a species tree $S$ and a set of gene trees $G$\citep{hide}:
\begin{enumerate}
    \item Pick a horizontal edge $H = ((u,v),(v,u))$ from $S$
    \item Pick a gene tree $G_i$ in $G$
    \item Decompose $G_i$ into it's set of quartets $\phi_i$
    \item Remove all quartets consistent with $S$ or previously explained from $\phi_i$
    \item Set $RS((u,v),\phi_i)$ to be the number of quartets in $\phi_i$ that support the edge $(u,v)$
    \item Set $NS((u,v),\phi_i)$ to be $RS((u,v),\phi_i)$ divided by $\lambda$, which is the total number of quartets in $S$ that are consistent with the edge $(u,v)$.
    \item The score for the edge $H$ for tree $G_i$ is $max\{NS((u,v),\phi_i),NS((v,u),\phi_i)\}$
    \item The total score for the edge $H$ is the sum of scores for each tree $G_i$
    \item This total score calculation is repeated for each horizontal edge $H_i$ in S, resulting in a list of edges, which is a complete description of the network.
\end{enumerate}
\subsection{Network Statistics}
All networks will be comprised of nodes representing \ac{otu}s and weighted edges represent the estimated amount of \ac{hgt} events between the two incident nodes.
As multiple set of networks can be computed for a single set of genera (using different sets of gene trees), bootstrap support for edges and confidence intervals on edge weights can also be calculated.
Given a network, with a set of nodes $V = \{V_0$ \dots $V_i\}$ of cardinality $N$ and a set of weighted edges (an unordered 2-tuple and weight) $T = \{((V_1,V_2),W_{1,2})$ \dots $((V_i,V_j),W_{i,j})\}$ with cardinality $E$ descriptive statistics can be computed as follows\citep{netstat}:
\begin{itemize}
    \item Total edge weight: sum of all edge weights in a network
    \item Average edge weight: sum of all edge weights divided by $N$
    \item Node Closeness Centrality:$ \frac{N-1}{\sum_v d(x,v)}$ where $d(x,y)$ is the length of the shortest path between node v and x.
    \item Node Associativity:$ \frac{j(j+1)(\overline{k}-\mu_q)}{2E\sigma^2_q}$ where $j$ is the excess degree of the node and $\overline{k}$ is the average excess degree of the node's neighbors and $\mu_q$ and $\sigma_q$ are the mean and standard variation of the excess degree distribution.
    \item Network Density:$ \frac{2(E-N+1)}{N(N-3)+2}$
    \item Node Clustering Coefficient:$ \frac{2e}{k(k-1)}$ where $k$ is the number of neighbors and $e$ is the number of edges between all neighbors.
    \item Network Diameter: The shortest path between the 2 furthest nodes in a network.
\end{itemize}
\printbibliography

